BCPNN and Sequence Learning

### 2018-09-18 document/draft.tex
Non-orthgonal sequences

\usepackage{graphicx}
\graphicspath{{images/}}

@@

\section{Introduction}

\subsection{Sequence Learning}
Section with a brief history of sequence learning in computational nueroscience. Key papers, order, structure.

The problem started with Lashley \cite{lashley1951problem} Hebb reference.

The other papers

\subsection{The BCPNN}
Explanation of what is the BCPNN, where does it come from, and some applications.

\begin{figure}[H]
\centering
\includegraphics[scale=0.30]{bcpnn_probabiltiies.pdf}
\caption{A schematic of how the BCPNN learning rules associates weights to different probability scenarios.}
\label{fig:bcpnn_probabilities}
\end{figure}

\begin{align}
p_i &= \frac{1}{N}\sum_{\mu=0}^{N} S_i^{\mu} \label{eq:prob_counting} \\  
p_{ij} &= \frac{1}{N} \sum_{\mu=0}^N S_i^{\mu} S_j^{\mu} \label{eq:prob_counting_co}
\end{align}

\begin{align}
\beta_i &=  \begin{cases}
       \log(p_i),&  p_i > 0,\\
       \log(\epsilon),& p_i = 0.0
    \end{cases} \label{eq:bcpnn_beta} \\
w &= \begin{cases}
       0 ,&  p_i=0 \text{ or } p_j = 0,\\
       \log(\epsilon),& p_{ij} = 0.0 \\
       \log(\frac{p_{ij}}{p_i p_j}),&  \text{otherwise}  
    \end{cases} \label{eq:bcpnn_w}
\end{align}

Where $\epsilon$ is a small number chosen out of convenience to avoid the fact that $log$ is not defined for $0$.

\subsection{A simple phenomenological theory of sequence recall}
Successful sequence recall dynamics can be described as three dynamical qualities and their interplay: fixation, erosion and biased transition.

\textbf{Fixation} is a general mechanism that fixes a state passing a certain point.

Examples of fixation. Attractor dynamics.

\textbf{Erosion} is a mechanisms that eventually suppresses the state after the system has dwell some time on it.

Examples of erosion 
Spike frequency adaptation \cite{roach2016memory}, feedback inhibition \cite{recanatesi2017memory}.

\textbf{Biased Transition}, after the state has been eroded the

Examples of biased transition
Biased by similiarty \cite{recanatesi2017memory}

\section{The BCPNN as a Sequence Learning Network}
First we can show that the BCPNN can be used do incremental learning \cite{sandberg2002bayesian}. And in particular there is a spike model that can be used to learn sequences \cite{tully2016spike}.

\subsection{Sequence Recall}
In order to test test the capabilities of the BCPNN neural network as a sequence learning mechanism we will start with a minimal model of it. The idea here is to show the minimal conditions under which the system can successfully reproduce sequential activity and isolate how the parameters and properties of the network interact with each other. After we are equipped with this knowledge we will add more parts to the model and this in turn will provide new capabilities that we will explore later.

Using the phenomenology presented before, we will explain how the structure of the network in figure \ref{fig:bcpnn_simple_network} and the dynamical equations of the system intertwine to achieve, fixation, erosion and biased transition altogether.

\begin{figure}[H]
\centering
\includegraphics[scale=1.40]{simple_BCPNN.pdf}
\caption{A simple BCPNN network with only one type of connectivity.}
\label{fig:bcpnn_simple_network}
\end{figure}

First the system achieves fixation by the means of the self-excitatory current depicted in red in figure \ref{fig:bcpnn_simple_network}. By itself this mechanism will fix all the patterns at the same time, that is why we need competitive selection.  To solve that problem we will use a winner-takes-all mechanism \cite{yuille1998winner} implemented in the form of equation \ref{eq:simple_bcpnn_max}. This equation ensures that at any point in time only the unit with the higher input current is activated.

After a particular unit is activated the adaptation current in equations \ref{eq:simple_bcpnn_adaptation} and \ref{eq:simple_bcpnn}  will be the mechanism responsible for the erosion of the pattern. Once a unit is activate for long enough and in the right parameter regime the adaptation current will surpass the self-excitatory current and the pattern will be suppressed. On this light, the time a particular unit remains activated is mostly dependent on the parameters that determine the dynamics of the adaptation and self-excitatory currents and the competitive balance between them. We will make this last relationship quantitative further down in this document.

Finally, this system would jump randomly among the states if it were not for a proper mechanism of biased transition. This is accomplished with differential inhibitory weights (illustrated in figure \ref{fig:bcpnn_simple_network} as different widths for the blue inhibitory connections) that become more and more inhibitory the farther two units are in the sequence. This ensures that once the adaptation currents for a unit becomes big enough the next unit which is the less inhibited one wins the competition and gets prompted to activation by the winner-takes-all mechanism.

\begin{align}
\tau_m \dfrac{ds_i}{dt} &= g_{beta}\beta_i + g_{w}\sum_{j} w_{ij} o_j  - g_a a_i - s_i \label{eq:simple_bcpnn} \\ 
o_i &=  \delta_{i, argmax(s)} \label{eq:simple_bcpnn_max} \\ 
\tau_a \dfrac{da_i}{dt} &= o_i - a_i \label{eq:simple_bcpnn_adaptation}
\end{align}

In order to illustrate how the dynamics of the system work together we show an example of a successful sequence recall in figure \ref{fig:bcpnn_simple_recall}. In the recall process we cue the first unit of the sequence by clamping it by a short period of time ($\sim 100ms$). We then let the system evolve on its own and, given the right combination of parameters, a sequence is effectively recalled if all the units that conform the pattern are activated in the expected order.

It is important to note that in this case we utilized a tailor-made connectivity matrix to clarify the relationship between the different component of the dynamics. Further down we will show how the same effect can be effectively achieved with a matrix that emerges from a self-organized learning process.

@@

If there was a previous activation of the unit the quantity $a_0$ will reflect how much the unit charged (and then discharged) before the current activation of the unit took place.

It is desirable to calculate how long does a unit $i$ becomes activated before the unit $j$ becomes activated. This is called persistent time $T_{persistence}$ and can be calculated from the equations above by making the assumption that the $T_{persistence}$ is far bigger than $\tau_s$, we also need to assume homogeneity in $\tau_a$.

\begin{align*}
\frac{T_{ij}}{\tau_a} &= \log \left(\frac{1}{1 - B_{ij}} \right) +\log \left( \frac{g_i}{g_j}a_i(0)f_i + (1 - a_j(0))f_j \right) 
\end{align*}

Where $B_{ij} = \frac{w_{jj} - w_{ij} + \beta_j - \beta_i}{g_j}$ is a measure of inertia or inverse of speed. The close to one it is, the longer the transition takes. Also, we see that $B_{ij}$ needs to be bigger than 0 for the transition to occur, from there we can derive the criteria that $g_j > \delta w + \delta \beta$. We also introduced the short hand $f_i = \frac{1}{1 - r_i}$ which is always bigger than 1 for this system and it represents delay effects introduced due to the capacitive nature of the uni.  We see that the presence of adaptive current in the unit that is active hastens the transition ($a_j(0) > 0$  makes $T_{ij}$ smaller by reducing the second term inside the second logarithm). On the other hand, the presence of adaptive current on the next unit ($a_i(0) > 0$) in the sequence slows the process by increasing the first term in the second logarithm.

@@

We can calculate from this the \textbf{backwards time} that is, once you passed through a particular unit.

\begin{figure}[H]
\centering
\includegraphics[scale=0.26]{simple_bcpnn_recall.pdf}
\caption{An instance of recall in the simple BCPNN neural network. a) Unit activity starting with the cue. b) the time course of the adaptation for each unit. c) the self-excitatory current minus the adaptation current, note that this quantity crossing the value of $w_{next}$ (depicted here with a dotted line) marks the transition point from one unit to the next. d) The connectivity matrix where we have included pointers to the three most important quantities $w_{self}$ for the self-excitatory weight, $w_{next}$ for the inhibitory connection to the next element and $w_{rest}$ for the rest of the connections.}
\label{fig:bcpnn_simple_recall}
\end{figure}

We will now proceed to characterize the properties of the simplified version of the BCPNN. We will do this in two steps, first we will explain the recall properties of the system and then we will proceed to describe the learning rule and its dynamics. One of the most important quantities of sequence recall network is the \textbf{persistence time.} of a a state. It turns out than in this simple system by determining the recall time with regards to certain parameters we can explain both how the network works and when the recall is carried out successfully.

\subsubsection{Persistence Time}
If we have system with only one type of connectivity, and winner-takes-all selectivity one state will be suppressed in favor of the other as soon as the support of the second state is bigger. In more detail, if we start with a system where the first unit is activated its own support will be $s_1 = g w_{self} - a(t)$ where a is the adaptation current which grows in time. This will continue as long as the first unit is activated and as a consequence the support for the unit will be decreasing. On the other hand the second unit is receiving a constant current to its support $s_2 = g w_{next}$, if this process continues by continuity there will be a point when the support of the first unit will be equal to that of the second:

\begin{align*}
s_{self} &= s_{next} \\
g_{\beta} \beta_{self} + g_w w_{self} - g_{a} (1 - e^{\frac{t}{\tau_a}}) &= g_{\beta} \beta_{next} + g_w w_{next}
\end{align*}
%g_{beta} \beta w_{self}

Where we have substituted the proper term for adaptation. We can solve for $t$ in the expression above to obtain the persistent time $T_{persistent}$:

\begin{align}
T_{persistence} = \tau_{a} \ln \left(\frac{g_a}{g_a - g_w (w_{self}  - w_{next}) - g_{beta} (\beta_{self} - \beta_{self})} \right) \label{eq:simple_bcpnn_persistence_time}
\end{align}

We notice that the equation behaves linearly with regards to $\tau_a$ and logarithmically with regards to $g_a$, $g_w$, $w{self}$ $w_{next}$. Even more importantly, certain combinations of the latter parameters will give raise to singularity points. This is important because close to this values the persistence time can be varied greatly with small variations in the respective parameters. In other words we have a wide dynamical variability in $T_{persistence}$ as in \cite{murray2017learning}. We proceed now to describe the scaling for each of the parameters in a more concise way and to compare the theory derived above with simulations.

\textbf{Adaptation current time constant} $\mathbf{\tau_{a}}$ \\
We explain here how the persistence time dynamics depend on $\tau_a$ the time constant of the adaptation current. A priori, the longer the adaptation time constant the longer it will take to the adaptation time current to erode the pattern. From equation \ref{eq:simple_bcpnn_persistence_time} we can observe that the relationship is linear which is exactly what we get in figure \ref{fig:simple_bcpnn_comparison} a).  Note however, that the slope is not very step, that is, we do not get a lot of variation on the persistence time from variations on $\tau_a$.

\textbf{Adaptation current gain $g_a$}
The adaptation time current is actually a limiting factor in the succesfull recall of a sequence. If the adaptation current is not big enough to overcome the difference between the self-excitatory current and the current of the next element then the system will get stuck forever in the same state.  We need therefore that the value of the adaptation gain $g_a$ to be bigger the difference in weights multiplied by the weight gain $ g_w (w_{self}  - w_{next})$ which is just the denominator of the equation \ref{eq:simple_bcpnn_persistence_time}. Once we have overcome this threshold the behavior becomes obvious, the bigger the gain of the adaptation current the faster this current will overcome the self-excitatory one, as a consequence the persistence time will be smaller. We illustrate this behavior in figure \ref{fig:simple_bcpnn_comparison} b). The singularity here just reflects the fact when the adaptation is barley enough to overcome the difference in currents it takes an exponential time to actually do it.

\begin{figure}[H]%
    \centering
    \subfigure[$\tau_a$]{{\includegraphics[width=5cm]{simple_bcpnn_tau_a.pdf} }}%
    \qquad
    \subfigure[$g_a$]{{\includegraphics[width=5cm]{simple_bcpnn_g_a.pdf} }}%
    \hfill
    \subfigure[$g_w$]{{\includegraphics[width=5cm]{simple_bcpnn_g_w.pdf} }}%
     \qquad
    \subfigure[$w_{next}$]{{\includegraphics[width=5cm]{simple_bcpnn_w_next.pdf} }}%
    \caption{Persistence time relationship with the parameters. a) We can appreciate that the persistence time grows linearly with $\tau_a$, the adaptation current time constant. b) Here we depict the logarithimic dependence of $T_{persistence}$ on $g_a$. c) The same dependence for $g_w$. d) We illustrate here the effects of making the weight differential bigger (note that $w_{self}=1$ in this plot). }
    \label{fig:simple_bcpnn_comparison}%
\end{figure}

\textbf{Weight Gain}
If we look at the denominator of equation \ref{eq:simple_bcpnn_persistence_time} we can interpret $g_w$ as an amplifier of the weight differential ($w_{self} - w_{next}$, this in turn is a proxy for the amount of current that the adaptation current has to overcome. In the light of this is not surprising the the persistence time increases with $g_w$. Moreover, once the weight differential gets amplified enough the adaptation current will take an exponential amount of time to overcome this difference given rise to a singularity as well.

\textbf{Next weight value}
In order to quantify the effects of varying $w_{next}$ we fixed the value of $w_{self}$ to $1$. With this perspective on mind, the farther $w_{next}$ is from 1 (the red line on the right) the biggest is the weight differential and using the same reasoning that we used for the effects of $g_w$ above we conclude increasing persistent time and a singularity when the difference becomes big enough (red line to the left).

\subsection{Sequence Learning}
\subsubsection{Off-line learning rule}

In order for the BCPNN to learn from temporal patterns a first step is to generalize the rules given in equations \ref{eq:prob_counting_co} and \ref{eq:prob_counting_co} to cope with continuous signal. We propose here the following equations:

\begin{align}
p_i = \frac{1}{T}\int_0^{T} S_i(t) dt  \label{eq:bcpnn_off_line_prob} \\
p_{ij} = \frac{1}{T}\int_0^{T} S_i(t) S_j(t) dt \label{eq:bcpnn_off_line_joint} 
\end{align}

Where  $T$ is the total presentation time. Note that this rules are a direct generalization of the discrete rules and can be reduced to them under the appropriate conditions, nevertheless they can deal directly deal with patterns of different duration and with incomplete activation in a way that is consistent with the normal rule. Once we have calculated the probabilities from the complete signal in the equations above we can use the $p_i$ and $p_ij$ quantities to estimate $w$ and $\beta$ using equations \ref{eq:bcpnn_w} and \ref{eq:bcpnn_beta}.

In figure \ref{fig:off_line_rule_signal} we illustrate how a collection of patterns presented in succession (left) look as continuous signals over time.

\begin{figure}[H]
    \centering
    \includegraphics[width=10.5cm]{off_line_rule_illustration_signal.pdf} 
	\label{fig:off_line_rule_signal}
	\caption{left: a collection of patterns presented to the neural network. right) The same set of patterns }
\end{figure}

In figure \ref{fig:off_line_rule_weights} we compare the both the weights $w$ and the co-activations $p_ij$ estimated from both the rule for discrete signals and the continuous signals. As we can appreciate the results are the same in both cases.

\begin{figure}[H]
    \centering
     \includegraphics[width=10.5cm]{off_line_rule_illustration_weight.pdf} %
	\label{fig:off_line_rule_weights}
	\caption{left) estimation of both co-activations and $w$ for the discrete presentation of patterns. right) estimation of both co-activations and $w$ for the continuous signal. }
\end{figure}

Finally if we assume that we are estimating the probabilities from signals whose information is delay and interpreted through a window of time

\begin{align}
z(t) &= \frac{1}{\tau_z } \int_{-\infty}^{t} s(\tau) e^{-\frac{t - \tau}{\tau_z}} d\tau \label{eq:flitering} 
\end{align}

\begin{figure}[H]
    \centering
  \includegraphics[width=10.5cm]{off_line_rule_illustration_filter.pdf} %
	\label{fig:off_line_rule_filter}
\end{figure}

An example of such filtering of the signal is illustrated on figure \ref{fig:off_line_rule_filter}. In the figure we can appreciate that the signal ``leaks' on time and as a consequence there is some overlap between patterns that are contiguous on time where before there was none. It follows from there that the co-activations and in consequence the connectivity $w$ will connect patterns that are close on time as illustrated on figure \ref{fig:off_line_rule_filters_weights}

\begin{figure}[H]
    \centering
    \includegraphics[width=10.5cm]{off_line_rule_illustration_weight_filtered.pdf}
	\label{fig:off_line_rule_filters_weights}
\end{figure}

\textbf{Training time}

\begin{figure}[H]
    \centering
    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_training_time_w.pdf} }}%
    \qquad
    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_training_time_p.pdf} }}%
    \caption{Results of training with different training times. a) The weights after training. b) probabilities after training.}
    \label{fig:off_line_learning_training_time}%
\end{figure}

\begin{figure}[H]
    \centering
    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_training_time_time.pdf} }}%
    \qquad
    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_training_time_diff.pdf} }}%
    \caption{Results of training with different training times. a) The weights after training. b) probabilities after training.}
    \label{fig:off_line_learning_training_time2}%
\end{figure}

$\mathbf{\tau_z}$

\begin{figure}[H]
    \centering
    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_tau_z_w.pdf} }}%
    \qquad
    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_tau_z_p.pdf} }}%
    \caption{Results of training with different values of $\tau_z$. a) The weights after training. b) probabilities after training.}
    \label{fig:off_line_learning_tau_z}%
\end{figure}

\begin{figure}[H]
    \centering
    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_tau_z_time.pdf} }}%
    \qquad
    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_tau_z_diff.pdf} }}%
    \caption{Results of training with different training times. a) The weights after training. b) probabilities after training.}
    \label{fig:off_line_learning_tau_z2}%
\end{figure}

\textbf{Epochs}

\begin{figure}[H]
    \centering
    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_epochs_w.pdf} }}%
    \qquad
    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_epochs_p.pdf} }}%
    \caption{Results of training with different number of epochs. a) The weights after training. b) probabilities after training.}
    \label{fig:off_line_learning_epochs}%
\end{figure}

\begin{figure}[H]
    \centering
    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_epochs_time.pdf} }}%
    \qquad
    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_epochs_diff.pdf} }}%
    \caption{Results of training with different training times. a) The weights after training. b) probabilities after training.}
    \label{fig:off_line_learning_epochs2}%
\end{figure}

\textbf{Inter Sequence Interval}

\begin{figure}[H]
    \centering
    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_ISI_w.pdf} }}%
    \qquad
    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_ISI_p.pdf} }}%
    \caption{Results of training with different values of the inter sequence interval. a) The weights after training. b) probabilities after training.}
    \label{fig:off_line_learning_ISI}%
\end{figure}

\begin{figure}[H]
    \centering
    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_ISI_time.pdf} }}%
    \qquad
    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_ISI_diff.pdf} }}%
    \caption{Results of training with different training times. a) The weights after training. b) probabilities after training.}
    \label{fig:off_line_learning_ISI2}%
\end{figure}

\textbf{Inter Pulse Interval}

\begin{figure}[H]
    \centering
    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_IPI_w.pdf} }}%
    \qquad
    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_IPI_p.pdf} }}%
    \caption{Results of training with different values of the inter pulse interval. a) The weights after training. b) probabilities after training.}
    \label{fig:off_line_learning_IPI}%
\end{figure}

\begin{figure}[H]
    \centering
    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_IPI_time.pdf} }}%
    \qquad
    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_IPI_diff.pdf} }}%
    \caption{Results of training with different training times. a) The weights after training. b) probabilities after training.}
    \label{fig:off_line_learning_IPI2}%
\end{figure}

\textbf{Resting Time}

\begin{figure}[H]
    \centering
    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_resting_w.pdf} }}%
    \qquad
    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_resting_p.pdf} }}%
    \caption{Results of training with different values of resting time. a) The weights after training. b) probabilities after training.}
    \label{fig:off_line_learning_resting}%
\end{figure}

\begin{figure}[H]
    \centering
    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_resting_time.pdf} }}%
    \qquad
    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_resting_diff.pdf} }}%
    \caption{Results of training with different training times. a) The weights after training. b) probabilities after training.}
    \label{fig:off_line_learning_resting2}%
\end{figure}

\textbf{Minicolumns}

\begin{figure}[H]
    \centering
    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_minicolumns_var_w.pdf} }}%
    \qquad
    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_minicolumns_var_p.pdf} }}%
    \caption{Results of training with different values of resting time. a) The weights after training. b) probabilities after training.}
    \label{fig:off_line_learning_minicolumns_var}%
\end{figure}

\begin{figure}[H]
    \centering
    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_minicolumns_var_time.pdf} }}%
    \qquad
    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_minicolumns_var_diff.pdf} }}%
    \caption{Results of training with different training times. a) The weights after training. b) probabilities after training.}
    \label{fig:off_line_learning_minicolumns_var2}%
\end{figure}

\textbf{Minicolumns - fixed sequence}

\begin{figure}[H]
    \centering
    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_minicolumns_fixed_w.pdf} }}%
    \qquad
    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_minicolumns_fixed_p.pdf} }}%
    \caption{Results of training with different values of resting time. a) The weights after training. b) probabilities after training.}
    \label{fig:off_line_learning_minicolumns_fixed}%
\end{figure}

\begin{figure}[H]
    \centering
    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_minicolumns_fixed_time.pdf} }}%
    \qquad
    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_minicolumns_fixed_diff.pdf} }}%
    \caption{Results of training with different values of resting time. a) The weights after training. b) probabilities after training.}
    \label{fig:off_line_learning_minicolumns_fixed2}%
\end{figure}

\textbf{Hypercolumns}

\begin{figure}[H]
    \centering
    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_hypercolumns_w.pdf} }}%
    \qquad
    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_hypercolumns_p.pdf} }}%
    \caption{Results of training with different values of resting time. a) The weights after training. b) probabilities after training.}
    \label{fig:off_line_learning_hypercolumns}%
\end{figure}

\begin{figure}[H]
    \centering
    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_hypercolumns_time.pdf} }}%
    \qquad
    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_hypercolumns_diff.pdf} }}%
    \caption{Results of training with different values of resting time. a) The weights after training. b) probabilities after training.}
    \label{fig:off_line_learning_hypercolumns2}%
\end{figure}

\subsubsection{On-line vs off-line filters}
Here we show the equivalence between the formulation of the filters as differential equations and the filters as convolutions. 
\begin{align*}
\tau_z \dfrac{dz}{dt} &= s(t) - z(t) \\
\tau_z \mathcal{F}\{\dfrac{dz}{dt}\} &= \mathcal{F}\{s(t)\} -  \mathcal{F}\{ z(t) \}  \text{ \qquad Applying the Fourier transform} \\
\tau_z i \nu \mathcal{F}\{ z(t) \} &=  \mathcal{F}\{s(t)\} -  \mathcal{F}\{ z(t) \}  \text{ \qquad Using the derivate rule} \\
\mathcal{F}\{ z(t) \}(1 + \tau_z i \nu) &= \mathcal{F}\{s(t)\} \\ 
\mathcal{F}\{ z(t) \} &= \mathcal{F}\{s(t)\} \left(\frac{1}{1 +\tau_z i \nu } \right) \text{ \qquad Factorizing and rearranging} \\ 
z(t) &= \mathcal{F}^{-1} \Big\{ \mathcal{F}\{s(t)\} \cdot \left(\frac{1}{1 +\tau_z i \nu } \right) \Big\} \text{ \qquad Inverse}  \\ 
z(t) &= \mathcal{F}^{-1} \Big\{ \mathcal{F}\{s(t)\} \Big\} \circledast \mathcal{F}^{-1} \Big\{  \frac{1}{1 +\tau_z i \nu } \Big\} \text{ \qquad Convolution}  \\ 
z(t) &= \frac{1}{\tau_z } s(t) \circledast \mathcal{F}^{-1} \Big\{  \frac{1}{\frac{1}{\tau_z } + i \nu } \Big\} \text{ \qquad Identity and rearranging}  \\ 
z(t) &= \frac{1}{\tau_z } s(t) \circledast e^{-\frac{t}{\tau_z}}u(t)  \text{ \qquad Heaviside function u(t)}  \\ 
z(t) &= \frac{1}{\tau_z } \int_{-\infty}^{\infty} s(\tau) e^{-\frac{t - \tau}{\tau_z}}u(t - \tau) d\tau  \text{ \qquad Definition}  \\ 
z(t) &= \frac{1}{\tau_z } \int_{-\infty}^{t} s(\tau) e^{-\frac{t - \tau}{\tau_z}} d\tau  \text{ \qquad Heaviside argument positive}
\end{align*}

\textbf{Analytical probabilities}

We can calculate the probability of a single unit being activated with the following formula: 
\begin{align*}
T_{total} *  p_i &= T_{activation} + \tau_z e^{\dfrac{T_{start} - T_{total}}  {\tau_z}} \left( e^{\dfrac{T_{activation}}{\tau_z}} - 1  \right)
\end{align*}

Where $T_{activation}$ is the the time that unit was activated during the course of the training. $T_{start}$ is the time at which the unit was activated, and it takes into account frontier effects because of the filter. Finally $T_{total}$ is the total training time.

We can also calculate the probability of the joint probability between two units with the more complicated formula. 
%T1 / tau1 + (Ts + T2) / tau2 - Tt / tau_p
\begin{align*}
p_{21} &= \frac{A  + B  - C}{T_{total}} \\
A & = M_1  \tau_{pre} \left( e^{\dfrac{-(T_s - T_1)}{\tau_{pre}}} - e^{\dfrac{-(T_s + T_2 - T_1)}{\tau_{pre}}}  \right) \\
B &= M_1 \tau_p \left(e^{\dfrac{T_1}{\tau_{pre}} + \dfrac{T_s}{\tau_{post}} - \dfrac{T_s + T_2}{\tau_p}} - e^{\dfrac{T_1}{\tau_{pre}} + \dfrac{T_s}{\tau_{post}} - \dfrac{T_s}{\tau_p}}   \right) \\
C &=  M_1 M_2 \tau_p \left(e^{\dfrac{T_1}{\tau_{pre}} + \dfrac{T_s + T_2}{\tau_{post}} -\dfrac{T_{total}}{\tau_p}} - e^{\dfrac{T_1}{\tau_{pre}} + \dfrac{T_s + T_2}{\tau_{post}} -\dfrac{T_S + T_2}{\tau_p}} \right)  \\
M_1 & =  1 - e^{-\dfrac{T_1}{\tau_{pre}}} \\ 
M_2 & = 1 - e^{-\dfrac{T_2}{\tau_{post}}} \\
\tau_p &=  \dfrac{\tau_{pre} \tau_{post}}{\tau_{pre} + \tau_{post}}
\end{align*}

Where $T_1$ is the time that the first unit is activated, $T_2$ is the time that the second unit remained activated, $T_s$ is the time at which the second unit become activated (If second unit is activated immediately after the first one then $T1=T_s$), finally $T_{total}$ is the total training time.

We can also calculate the joint probability of a unit with itself with the following formula

\begin{align*}
p_{11} &= \frac{A + B}{T_{total}} \\
A &= T_1 -  M_{pre} \tau_{pre} - M_{post} \tau_{post} + M_p \tau_p  \\
B &= \tau_p M_{pre} M_{post} \left( 1 - e^{-\dfrac{T_{total} - T_1}{\tau_p}} \right) \\ 
M_{pre} &= 1 - e^{-\dfrac{T_1}{\tau_{pre}}} \\ 
M_{post} &= 1 - e^{-\dfrac{T_1}{\tau_{post}}} \\
M_p &= 1 - e^{-\dfrac{T_1}{\tau_{p}}} \\
\tau_p &= \dfrac{\tau_{pre} \tau_{post}}{\tau_{pre} + \tau_{post}}
\end{align*}

\subsubsection{On-line rule}
Once we know the dynamics of the system given a certain matrix the natural question to consider is whether we can learn the weight matrix. As described in \cite{sandberg2002bayesian} with the help of traces we can add on-line learning capabilities to the BCPNN neural network.

\begin{align}
\tau_z \dfrac{dz_i}{dt} &= o_i - z_i \label{eq:traces}\\
\tau_p \dfrac{dp_i}{dt} &= z_i - p_i  \label{eq:traces_probability} \\  
\tau_p \dfrac{dp_{ij}}{dt} &= z_i z_j - p_{ij} \\
w_{ij} &= \log(\frac{p_{ij}}{p_i p_j}) \label{eq:bcpnn_weight_update} \\
\beta_i &= \log(p_i) 
\end{align}

We can understand the Bayesian nature of this learning rule by focusing our attention in equation \ref{eq:bcpnn_weight_update}. The argument of the logarithm weights co-activations of the two units against the base rate activation of the units multiplied by each other. When that two units are activated at the same time the co-activation $p_{ij}$ of the units increases, if this is bigger than the  base rate of the activation of each unit multiplied ($p_i p_j$) then the weight increases, otherwise it decreases. We translate this to a problem of linking on time by using the z-traces as described in equation \ref{eq:traces} and illustrated in equation \ref{fig:traces_example}. If two units are close together in time there will be co-activation of the traces (illustrated in red) within a time scale of size $\tau_z$. Finally  we use a second low-pass filter  with longer time constant to preserve the learning for a longer time-scale $\tau_p$.

@@

\textbf{The training protocol:} \\
Here we describe the training protocol. The first to note is that we train the neural network by clamping a given set of units in the order of a given sequence for a given time. More specifically a training protocol is characterized by the following quantities. First there is the \textbf{training time} for a given element of the sequence, that is, the time that element remains activated. It may well be that the an element of the sequences does not follow the other immediately, in order to account for this possibility we have the \textbf{inter pulse interval (IPS)} which is just the time elapsed between a pattern and its successor. It is also the case that a given sequence of elements is often presented more than once, we call \textbf{epoch} to a particular instance of a sequence pretension. Finally, we usually leave some time between each of the epochs during the presentation of the training protocol, this is know as the \textbf{inter sequence interval (ISI)}. We present a schematic of the training protocol in figure \ref{fig:training_protocol}.

\begin{figure}[H]
\centering
\includegraphics[scale=1.40]{protocol.pdf}
\caption{The training protocol. IPI stands for inter pulse interval and ISI for inter sequence interval. Explanations in the text.}
\label{fig:training_protocol}
\end{figure}

>>>>

\usepackage{graphicx}
\graphicspath{{images2/}}

@@

\section{Results}
\begin{figure}[H]
\centering
\includegraphics[scale=1.0]{diagram.pdf}
\caption{A diagram }
\label{fig:networks_scheme}
\end{figure}

\begin{align}
\tau_s \dfrac{ds_i}{dt} &= g_{beta}\beta_i + \sum_{j} w_{ij} o_j  - g_a a_i - s_i  + \xi(t) \label{eq:simple_bcpnn} \\ 
o_i &=  \delta_{i, argmax(s)} \label{eq:simple_bcpnn_max} \\ 
\tau_a \dfrac{da_i}{dt} &= o_i - a_i \label{eq:simple_bcpnn_adaptation}
\end{align}

\subsection{The system can recall sequences}
\section{Results}
\begin{figure}[H]
\centering
\includegraphics[scale=0.25]{simple_bcpnn_recall.pdf}
\caption{An example of recall}
\label{fig:recall}
\end{figure}

@@

\subsection{Persistent time}
\begin{figure}[H]
\centering
\includegraphics[scale=0.25]{persistent_times.pdf}
\caption{A characterization of persistent times}
\label{fig:per_time}
\end{figure}

@@

Where $B_{ij} = \frac{w_{jj} - w_{ij} + \beta_j - \beta_i}{g_j}$ is a measure of inertia or inverse of speed. The close to one it is, the longer the transition takes. Also, we see that $B_{ij}$ needs to be bigger than 0 for the transition to occur, from there we can derive the criteria that $g_j > \delta w + \delta \beta$. We also introduced the short hand $f_i = \frac{1}{1 - r_i}$ which is always bigger than 1 for this system and it represents delay effects introduced due to the capacitive nature of the unit.  We see that the presence of adaptive current in the unit that is active hastens the transition ($a_j(0) > 0$  makes $T_{ij}$ smaller by reducing the second term inside the second logarithm). On the other hand, the presence of adaptive current on the next unit ($a_i(0) > 0$) in the sequence slows the process by increasing the first term in the second logarithm.

\subsection{Learning}
\begin{figure}[H]
\centering
\includegraphics[scale=1.40]{protocol.pdf}
\caption{The training protocol. IPI stands for inter pulse interval and ISI for inter sequence interval. Explanations in the text.}
\label{fig:training_protocol}
\end{figure}

@@

\begin{figure}[H]
\centering
\includegraphics[scale=0.15]{recall_example.pdf}
\caption{A recall example}
\label{fig:recall_example}
\end{figure}

\begin{figure}[H]
\centering
\includegraphics[scale=0.15]{training.pdf}
\caption{A characterization of the connectivity as a function of the protocol}
\label{fig:training}
\end{figure}

\subsection{Noise}

\begin{figure}[H]
\centering
\includegraphics[scale=0.30]{noise_diagram.pdf}
\caption{A schematic of the noise}
\label{fig:noise_scheme}
\end{figure}

\begin{figure}[H]
\centering
\includegraphics[scale=0.20]{noise_robustness.pdf}
\caption{A characterization of the noise}
\label{fig:robustness}
\end{figure}

\subsection{Representations}
\begin{figure}[H]
\centering
\includegraphics[scale=0.90]{overlap_diagram.pdf}
\caption{Overlap diagram}
\label{fig:representation_scheme}
\end{figure}

\begin{figure}[H]
\centering
\includegraphics[scale=0.20]{representations.pdf}
\caption{A characterization of the different overlap conditions}
\label{fig:representations}
\end{figure}

\subsection{Non-homogeneous training conditions}
